Interval-Valued Quantum Probabilities: Contextuality and the Born Rule

### 2017-07-06 prop_letter.tex
sec 4

The argument refers to the expectation value of these
observables. In the case of QIVPM this is defined as follows.

\begin{definition}Given a Hilbert space of $d\ge3$, for every
  QIVPM\nb$\bar{\mu}:\events\rightarrow\mathscr{I}$, any
  observable\nb$\mathbf{O}$ with spectral decomposition
  $\mathbf{O}=\sum_{i=1}^{N}\lambda_{i}P_{i}$, and $\events'$ is
  minimal sub-event space containing
  $\left\{ P_{i}\right\} _{i=1}^{N}$, we have
\begin{equation}
\expval{\mathbf{O}}_{\bar{\mu}}=\left[\min_{\rho\in\coreBorn\left(\bar{\mu},\events'\right)}\expval{\mathbf{O}}_{\rho},\max_{\rho\in\coreBorn\left(\bar{\mu},\events'\right)}\expval{\mathbf{O}}_{\rho}\right]\,.\label{eq:quantum-interval-expectation-core}
\end{equation}
\end{definition}

The expectation value is intuitively between two possible outcomes
which lie between the minimum and maximum bounds of the given
probability intervals, and this definition also consistent with the classical
Choquet integral on commuting sub-event space by adopting the classical
results\nb\cite{Rosenmuller1971,GilboaSchmeidler1994,Grabisch2016}.
If the QIVPM is cryptodeterministic, this
interval collapses to a singleton $\left\{ \lambda\right\} $, where
$\lambda$ is an eigenvalue of $\mathbf{O}$. If we define the
multiplication of two singleton sets\nb$\left\{ \lambda_{1}\right\} $
and $\left\{ \lambda_{2}\right\} $ to be
$\left\{ \lambda_{1}\lambda_{2}\right\} $, then we have
\begin{equation}
\prod_{j=1}^{N}\expval{\mathbf{O}_{j}}_{\bar{\mu}}=\expval{\prod_{j=1}^{N}\mathbf{O}_{j}}_{\bar{\mu}}\label{eq:Kochen-Specker-Rules}
\end{equation}
for any sequence of commuting observables\nb$\left\{ \mathbf{O}_{j}\right\} _{j=1}^{N}$.

Since the expectation value\nb$\expval{\cdot}_{\bar{\mu}}$ satisfying
these two properties, the observables in the Mermin-Peres square must
have the the expectation value\nb$\left\{ 1\right\} $ or $\left\{ -1\right\} $.
In each row and column, operators commutes, and each operator is the
product of the two others, except in the third column $\left(\sigma_{z}\otimes\sigma_{z}\right)\left(\sigma_{x}\otimes\sigma_{x}\right)=-\sigma_{y}\otimes\sigma_{y}$.
Therefore, it is clear impossible to find a IVPM\nb$\bar{\mu}$ such
that $\langle \mathbf{O}_{ij}\rangle_{\bar \mu}$
%\begin{equation}
%\begin{aligned}\langle\mathbb{1} & \otimes\sigma_{z}\rangle_{\bar{\mu}} & \langle\sigma_{z} & \otimes\mathbb{1}\rangle_{\bar{\mu}} & \langle\sigma_{z} %& \otimes\sigma_{z}\rangle_{\bar{\mu}}\\
%\langle\sigma_{x} & \otimes\mathbb{1}\rangle_{\bar{\mu}} & \langle\mathbb{1} & \otimes\sigma_{x}\rangle_{\bar{\mu}} & \langle\sigma_{x} & %\otimes\sigma_{x}\rangle_{\bar{\mu}}\\
%\langle\sigma_{x} & \otimes\sigma_{z}\rangle_{\bar{\mu}} & \langle\sigma_{z} & \otimes\sigma_{x}\rangle_{\bar{\mu}} & \langle\sigma_{y} & %\otimes\sigma_{y}\rangle_{\bar{\mu}}
%\end{aligned}
%\label{eq:MerminSquare-values}
%\end{equation}
satisfying Eq.~(\ref{eq:Kochen-Specker-Rules}).

According to Mermin's comment\nb\citep{Mermin1999}, 
\begin{quote}
...although the outcomes deduced from such imperfect measurements
will occasionally differ dramatically from those allowed in the ideal
case, if the misalignment is very slight, the statistical distribution
of outcomes will differ only slightly from the ideal case... 
\end{quote}
if the imperfection of the measurement is small enough, the finitely
precise measurement should be close to idealized measurement, and
cannot nullify the Kochen-Specker theorem. This idea can be modeled
precisely by considering the following kind of QIVPMs.

@@

For example, any cryptodeterministic QIVPM is $0$-deterministic.
As another example, a $\frac{1}{3}$-deterministic QIVPM can be given
by $\bar{\mu}\left(P\right)=\left[x/3,x/3\right]$, where $P$ projects
onto a $x$-dimensional subspace in a three-dimensional Hilbert space.
By utilizing $\delta$-deterministic QIVPMs, we can have a Kochen-Specker
theorem with finite precision measurement.

\begin{thm}\label{cor:Kochen-Specker-IVPM}Given a Hilbert space
of dimension\nb$d\ge3$, there is no $\delta$-deterministic QIVPM
for $\delta<\frac{1}{3}$.\end{thm}

\begin{proof}Suppose there is a $\delta$-deterministic QIVPM\nb$\bar{\mu}:\events\rightarrow\mathscr{I}$,
we can define a cryptodeterministic QIVPM\nb$\bar{\mu}^{\textrm{D}}:\events\rightarrow\left\{ \imposs,\necess\right\} $
by 
\begin{equation}
\bar{\mu}^{\textrm{D}}\left(P\right)=\begin{cases}
\imposs\,, & \textrm{ if }\bar{\mu}\left(P\right)\subseteq\left[0,\delta\right]\:;\\
\necess\,, & \textrm{ if }\bar{\mu}\left(P\right)\subseteq\left[1-\delta,1\right]\:.
\end{cases}
\end{equation}
However, there is no cryptodeterministic QIVPM as we discussed according
to the Mermin-Peres square, so that there is no $\delta$-deterministic
QIVPM.

The essential part to verify $\bar{\mu}^{\textrm{D}}$ is a cryptodeterministic
QIVPM is to verify a special case of Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion}),\begin{subequations}
\begin{eqnarray}
 &  & \bar{\mu}^{\textrm{D}}\left(P_{0}\right)=\bar{\mu}^{\textrm{D}}\left(P_{1}\right)=\imposs\\
 & \Rightarrow & \bar{\mu}^{\textrm{D}}\left(P_{0}+P_{1}\right)=\imposs
\end{eqnarray}
\end{subequations}for orthogonal operators\nb$P_{0}$ and $P_{1}$.

When $\bar{\mu}^{\textrm{D}}\left(P_{0}\right)=\bar{\mu}^{\textrm{D}}\left(P_{1}\right)=\imposs$,
we have both $\bar{\mu}\left(P_{0}\right)$ and $\bar{\mu}\left(P_{1}\right)\subseteq\left[0,\delta\right]$
which implies $\bar{\mu}\left(P_{0}+P_{1}\right)\subseteq\left[0,2\delta\right]$.
Since $\delta<\frac{1}{3}\Rightarrow2\delta<1-\delta$, $\bar{\mu}\left(P_{0}+P_{1}\right)$
cannot be a subset of $\left[1-\delta,1\right]$, but $\bar{\mu}\left(P_{0}+P_{1}\right)$
is a subset of either $\left[0,\delta\right]$ or $\left[1-\delta,1\right]$.
Hence, $\bar{\mu}\left(P_{0}+P_{1}\right)$ must be a subset of $\left[0,\delta\right]$,
which implies $\bar{\mu}^{\textrm{D}}\left(P_{0}+P_{1}\right)=\imposs$.

By discussing the remaining cases, we can easily verify that $\bar{\mu}^{\textrm{D}}$
is a cryptodeterministic QIVPM.\end{proof}

>>>>

\noindent The argument refers to the expectation value of these
observables. In the case of QIVPM this is defined as follows.

\begin{definition}Given a Hilbert space of $d\ge3$, for every
  QIVPM\nb$\bar{\mu}:\events\rightarrow\mathscr{I}$, any
  observable\nb$\mathbf{O}$ with spectral decomposition
  $\mathbf{O}=\sum_{i=1}^{N}\lambda_{i}P_{i}$, and any minimal
  sub-event space $\events'$ containing
  $\left\{ P_{i}\right\} _{i=1}^{N}$, we have
\begin{equation}
\expval{\mathbf{O}}_{\bar{\mu}}=\left[\min_{\rho\in\coreBorn\left(\bar{\mu},\events'\right)}\expval{\mathbf{O}}_{\rho},\max_{\rho\in\coreBorn\left(\bar{\mu},\events'\right)}\expval{\mathbf{O}}_{\rho}\right]\,.\label{eq:quantum-interval-expectation-core}
\end{equation}
\end{definition}

The expectation value is intuitively between two possible outcomes
which lie between the minimum and maximum bounds of the given
probability intervals. (This definition is consistent with the
classical Choquet integral on commuting sub-event spaces by adopting
the classical
results\nb\cite{Rosenmuller1971,GilboaSchmeidler1994,Grabisch2016}.)
If the QIVPM is cryptodeterministic, this interval collapses to a
singleton $\left\{ \lambda\right\} $, where $\lambda$ is an eigenvalue
of~$\mathbf{O}$. If we define the multiplication of two singleton
sets\nb$\left\{ \lambda_{1}\right\} $ and
$\left\{ \lambda_{2}\right\} $ to be
$\left\{ \lambda_{1}\lambda_{2}\right\} $, then we have
\begin{equation}
\prod_{j=1}^{N}\expval{\mathbf{O}_{j}}_{\bar{\mu}}=\expval{\prod_{j=1}^{N}\mathbf{O}_{j}}_{\bar{\mu}}\label{eq:Kochen-Specker-Rules}
\end{equation}
for any sequence of commuting observables\nb$\left\{ \mathbf{O}_{j}\right\} _{j=1}^{N}$.

Since the expectation value\nb$\expval{\cdot}_{\bar{\mu}}$ satisfies
these two properties, the observables in the Mermin-Peres square must
have the the expectation value\nb$\left\{ 1\right\} $ or
$\left\{ -1\right\} $.  In each row and column, the operators commute,
and each operator is the product of the two others, except in the
third column
$\left(\sigma_{z}\otimes\sigma_{z}\right)\left(\sigma_{x}\otimes\sigma_{x}\right)=-\sigma_{y}\otimes\sigma_{y}$.
Therefore, it is clearly impossible to find a QIVPM\nb$\bar{\mu}$ such
that $\langle \mathbf{O}_{ij}\rangle_{\bar \mu}$
%\begin{equation}
%\begin{aligned}\langle\mathbb{1} & \otimes\sigma_{z}\rangle_{\bar{\mu}} & \langle\sigma_{z} & \otimes\mathbb{1}\rangle_{\bar{\mu}} & \langle\sigma_{z} %& \otimes\sigma_{z}\rangle_{\bar{\mu}}\\
%\langle\sigma_{x} & \otimes\mathbb{1}\rangle_{\bar{\mu}} & \langle\mathbb{1} & \otimes\sigma_{x}\rangle_{\bar{\mu}} & \langle\sigma_{x} & %\otimes\sigma_{x}\rangle_{\bar{\mu}}\\
%\langle\sigma_{x} & \otimes\sigma_{z}\rangle_{\bar{\mu}} & \langle\sigma_{z} & \otimes\sigma_{x}\rangle_{\bar{\mu}} & \langle\sigma_{y} & %\otimes\sigma_{y}\rangle_{\bar{\mu}}
%\end{aligned}
%\label{eq:MerminSquare-values}
%\end{equation}
satisfies Eq.~(\ref{eq:Kochen-Specker-Rules}).

The question now is what happens if we relax the perfect measurement
assumption that entails that the intervals are precisely $[0,0]$ and
$[1,1]$. What if the measurement imperfections occasionally cause an
ideally impossible event to be detected or an ideally certain event to
be not be detected. Mermin argues in this case that:
\begin{quote}
  \ldots although the outcomes deduced from such imperfect
  measurements will occasionally differ dramatically from those
  allowed in the ideal case, if the misalignment is very slight, the
  statistical distribution of outcomes will differ only slightly from
  the ideal case~\citep{Mermin1999}.
\end{quote}

We can formally test this hypothesis by considering the notion of
$\delta$-deterministic QIVPM below.

@@

A cryptodeterministic QIVPM is $0$-deterministic.  As $\delta$ gets
bigger, the QIVPM allows more and more indeterminate behavior. A
$\frac{1}{3}$-deterministic QIVPM can be given by
$\bar{\mu}\left(P\right)=\left[x/3,x/3\right]$, where $P$ projects
onto a $x$-dimensional subspace in a three-dimensional Hilbert space.

We now prove that for small enough $\delta$, the Kochen-Specker is
still valid, i.e., if the measurement imperfections are small enough,
we cannot nullify the Kochen-Specker theorem as
Meyer~\cite{PhysRevLett.83.3751} argued.

\begin{thm}\label{cor:Kochen-Specker-IVPM} Given a Hilbert space
  of dimension\nb$d\ge3$, there is no $\delta$-deterministic QIVPM for
  $\delta<\frac{1}{3}$.\end{thm}

\begin{proof} Suppose there is a $\delta$-deterministic
  QIVPM\nb$\bar{\mu}:\events\rightarrow\mathscr{I}$, we claim that the
  following QIVPM
  \nb$\bar{\mu}^{\textrm{D}}:\events\rightarrow\left\{
    \imposs,\necess\right\} $ is cryptodeterministic:
\begin{equation}
\bar{\mu}^{\textrm{D}}\left(P\right)=\begin{cases}
\imposs\,, & \textrm{ if }\bar{\mu}\left(P\right)\subseteq\left[0,\delta\right]\:;\\
\necess\,, & \textrm{ if }\bar{\mu}\left(P\right)\subseteq\left[1-\delta,1\right]\:.
\end{cases}
\end{equation}
However, since we know that there is no cryptodeterministic QIVPM,
$\bar{\mu}^{\textrm{D}}$ cannot exist and hence $\bar{\mu}$ cannot
exist.

It remains to verify $\bar{\mu}^{\textrm{D}}$ is a cryptodeterministic
QIVPM. The main part of that verification is to check:
\begin{subequations}
\begin{eqnarray}
 &  & \bar{\mu}^{\textrm{D}}\left(P_{0}\right)=\bar{\mu}^{\textrm{D}}\left(P_{1}\right)=\imposs\\
 & \Rightarrow & \bar{\mu}^{\textrm{D}}\left(P_{0}+P_{1}\right)=\imposs
\end{eqnarray}
\end{subequations}for orthogonal operators\nb$P_{0}$ and $P_{1}$.

When $\bar{\mu}^{\textrm{D}}\left(P_{0}\right)=\bar{\mu}^{\textrm{D}}\left(P_{1}\right)=\imposs$,
we have both $\bar{\mu}\left(P_{0}\right)$ and $\bar{\mu}\left(P_{1}\right)\subseteq\left[0,\delta\right]$
which implies $\bar{\mu}\left(P_{0}+P_{1}\right)\subseteq\left[0,2\delta\right]$.
Since $\delta<\frac{1}{3}\Rightarrow2\delta<1-\delta$, $\bar{\mu}\left(P_{0}+P_{1}\right)$
cannot be a subset of $\left[1-\delta,1\right]$, but $\bar{\mu}\left(P_{0}+P_{1}\right)$
is a subset of either $\left[0,\delta\right]$ or $\left[1-\delta,1\right]$.
Hence, $\bar{\mu}\left(P_{0}+P_{1}\right)$ must be a subset of $\left[0,\delta\right]$,
which implies $\bar{\mu}^{\textrm{D}}\left(P_{0}+P_{1}\right)=\imposs$. 
\end{proof}
